 ---
title: "Greenwood Furnace Forest Community Analysis"
output: html_document
---

One of the most striking features of the terrestrial ecosystems of our beloved planet is how drastic are the changes you can see in plant communities as you roam across an environmental gradient. Changes in moisture, elevation, latitude, cardinal orientation, slope, seasonality, and solar radiation – among a myriad of biotic factors such as competition and mutualisms – drive changes in community structure and composition.

Vegetation sampling has a long tradition and is a core component of Plant Community Ecology. As heterotrophs we depend completely on plant communities, and thus, we have tried hard to understand what drives their structure and differences on the globe.

We can think of three realms of factors that affect the membership of tree species in a given site or community:

	Ultimate factors – evolutionary constrains, biogeography, and dispersal
	Proximate factors – biotic interactions (both positive and negative)
	Chance – randomness and stochasticity in births and deaths

 As we walked in Greenwood Furnace forest we saw changes and patterns that caught our eye. Some of the patterns may be due to random and historical forces; others may have an ecological/biological basis. 


Question 1: Are there any clear-cut communities as a function of elevation and/or aspect?


Lets start by making a Non Metric Multidimensional Scaling Ordination or NMDS (more info here http://ordination.okstate.edu/overview.htm ) for the trees using the number of trees of each species found in each plot:

```{r}
library(vegan) ## install library vegan first if not in your R version
library(MASS) ## install library MASS first if not in your R version

trees<-read.csv(file.choose(), head = TRUE, row.names = 1)
m<-metaMDS(trees, distance="chao", k=2, try=100, trymax=10000)

treat=c(rep("south.low",3),rep("south.high",3),rep("north.high",3),rep("north.low",3))
ordiplot(m, type="n")
orditorp(m,display="species",col="black",air=0.1, cex=.7)
orditorp(m,display="sites", air=0.01,cex=1)
ordiellipse(m,groups=treat, kind = "se", conf = 0.95, draw = "lines", lwd = 3, lty = 2,  col = c(rep("blueviolet"), rep("red"), rep("turquoise3"), rep("green")), w = NULL, label=FALSE)


```

Since we have the option of looking at a community not only by number of individuals, but by size, we will redo the tree ordination this time using the sum of the basal area per species

```{r}
library(vegan)
library(MASS)
treedbh<-read.csv(file.choose(), head = TRUE, row.names = 1)
m<-metaMDS(treedbh, distance="chao", k=2, try=100, trymax=10000)

treat=c(rep("south.low",3),rep("south.high",3),rep("north.high",3),rep("north.low",3))
ordiplot(m, type="n")
orditorp(m,display="species",col="black",air=0.1, cex=.7)
orditorp(m,display="sites", air=0.01,cex=1)
ordiellipse(m,groups=treat, kind = "se", conf = 0.95, draw = "lines", lwd = 3, lty = 2,  col = c(rep("blueviolet"), rep("red"), rep("turquoise3"), rep("green")), w = NULL, label=FALSE)



```


Question 2: What is the main driver of community differences in the Greenwood Furnace forests?



Now lets make an ordination of the seedlings:

```{r}
seedlings<-read.csv(file.choose(), head = TRUE, row.names = 1)
m<-metaMDS(seedlings, distance="chao", k=2, try=100, trymax=10000)

treat=c(rep("south.low",3),rep("south.high",3),rep("north.high",3),rep("north.low",3))
ordiplot(m, type="n")
orditorp(m,display="species",col="black",air=0.1, cex=.7)
orditorp(m,display="sites", air=0.01,cex=1)
ordiellipse(m,groups=treat, kind = "se", conf = 0.95, draw = "lines", lwd = 3, lty = 2,  col = c(rep("blueviolet"),  rep("red"), rep("turquoise3"),rep("green")), w = NULL, label=FALSE)


```




Question 3: How does these drivers affect different demographic classes of trees (seedlings, adults)?





Now, you will create your own "r chunks" to examine some of the variables in "univariate fashion".


Question 4: how does elevation and aspect affect tree species richness? Use the following chunk to construct a one-way analysis using the supplied file "tree_richness.csv". Also include a figure (boxplots) summarizing the data.
```{r}

tree_richness<-read.csv("tree_richness.csv", header = TRUE)

attach(tree_richness)
names(tree_richness)

boxplot(richness~Asp*Elev, data=tree_richness,         col=(c("red","orange","khaki3", "khaki")),
        main="tree richness", xlab="factors", cex.axis = 0.7)


model<- glm(richness~Asp*Elev, family = poisson(link="log"))

summary(model)
anova(model)


```




Question 5: How does elevation and aspect affect seedling species density? Use the following chunk to construct a one-way analysis using the supplied file "tree_richness.csv". Also include a figure (boxplots) summarizing the data. You may choose to conduct a separate analysis and figure for aspect and elevation.




```


Question 6: Is the forest changing?
For this you can look at which species of tree are recruitment-limited. This means calculating the difference – for each target species in each plot – between the number of seedlings and the number of adult trees (seedlings-sapplings per species per plot). You can thus create responses that revolve around zero, or that are positive or negative. Positive responses mean that there are more seedlings/saplings than adult trees. Average differences near to zero will mean no change. Negative averages suggest species losing dominance. Report a figure summarizing your results.

In this chunck you can plot such differences for each of the tree species. You will have to first create a new data table combining, for each plot, ac column with the number of seedlings, and the number of adult trees to obtain the difference.

```{r}

tree_richness<-read.csv("changing_forest.csv", header = TRUE)
library(car)
attach(tree_richness)
names(tree_richness)

boxplot(Difference~Tree.Species, data=tree_richness,         col=(c("red","orange","khaki3", "khaki")),
        main="tree richness", xlab="factors", cex.axis = 0.7)


model<- glm(Difference~Tree.Species)

Anova(model)


```



Question 7: Can you think of a different one-way analysis with this data set that we haven't examined so far? Explain how would you do it.
