# Tomasssss #


title: "Thistle_Seeds_&_Roads"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown


1) First setup your working directory under the "Session" Menu. Select the file were you have your data file.

Then we will import our data set as a csv text file using these commands:


```{r}
thistle <- read.csv("thistle.csv", header=TRUE)
attach(thistle)
names(thistle)
```



2) Before doing anything, we will check the distribution of the two response variables (distance, angle) we want to analyze, and test for their normality.

distance:
```{r}

hist(thistle$distance_m, main = "Dispersal distance", col = "lightblue", xlab = "frequency")
shapiro.test(thistle$distance_m)
```

angle:
```{r}

hist(thistle$angle, main = "angle", col = "lightblue", xlab = "frequency")
```


##Question set 1: Do we need any data transformations?##

If so, do them here by creating an "r chunk". Include a new histogram and a normality test for each case - was data normally distibuted after transformation?

#your answer here#


Adapt the following command lines to transform any response variable:

```{r}

thistle$log_distance_m <- log(thistle$distance_m + 1)

head(thistle)

hist(thistle$log_distance_m, main = "Histogram of distance", col = "lightblue", xlab = "frequency")

shapiro.test(thistle$log_distance_m)

```






3) Now we can make some plots:

distance:
```{r}
boxplot(distance_m~Road*speed_limit, data=thistle, notch=TRUE, 
        col=(c("red","blue", "yellow")),
        main="dispersal distance", xlab="treatments", cex.axis = 0.6)

```


angle:
```{r}


```



##Question 2: Does the graphs suggest there are differences between treatments? Discuss the case of each response:


#your answer here#



4) A formal test will analyze all the factors at once, including their interaction with a Generalized Linear Mixed Model:


first install these libraries (if your R doesn't have them - use the #install.packages("MASS")# for example, etc.) and open them
```{r}
library(MASS)
library(lme4)
library(car)

```



Now we will specify our model:

```{r}

modeldistance<-glm(log_distance_m~Road+speed_limit,  data=thistle)
summary(modeldistance)

plot(modeldistance)

```


See the results in form of a Type II ANOVA:
```{r}

Anova(modeldistance)
library(visreg)
visreg(modeldistance)
```





Discuss whether or not roads and cars are affecting the dispersal distance of thistle seeds. Base your conclusions on the significance tests for each parameter.  

#your answer here#

